Supplementary Materials

### 2023-07-02 manuscript/manuscript.tex
minor ms changes

%\title{Towards a Complete Model of the Negative Selection Process in Humans}
\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}

@@

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. Negative selection (i.e. purifying selection) works
to eliminate such deleterious mutations from the population, which is essential
to maintaining the integrity of information in the genome needed to maintain a
lineage. Mutations under negative selection are kept at low frequencies
(Haldane?), contributing to human traits (XXX) and diseases. Moreover, the
negative selection process determines sequence conservation across phylogenetic
timescales, which is evident in genome-wide cross-species comparisons
\parencite{Siepel2005-wh}. As with most selection in the genome, the negative
selection process perturbs the allele frequencies of neighboring linked
variation. This reduces genetic variability around conserved segments and
creates a large-scale spatial ``linked selection signal" in diversity alone the
genome. Mutation and negative selection's contribution to this signal is
mediated by the deleterious mutation rate, the strength of selection in
conserved segments, and the spatial distribution of recombination rates and
conserved segments. Since genome-wide recombination maps and putatively
conserved segments are available for many species, researchers have combined
these with population genetic theory to estimate the mutation rate and
selection coefficients of deleterious mutations, as well as the overall
reduction in genetic variability due to the negative selection process.

@@

In this work, we extend another class of linked selection models that derive
from quantitative genetics to address limitations of the classic BGS model.
These models consider how polygenic fitness variation increases the rate of
stochastic allele frequency change as alleles become randomly linked to fitness
backgrounds. While these models can theoretically accommodate polygenic fitness
variation from any source as long as its rate of change is not too rapid, we
focus specifically on a deleterious-mutations-only model from Santiago and
Caballero (2016). This model is identical to the classic BGS model when
selection against deleterious mutations is strong, but it also correctly
predicts the reduction in diversity and substitutions rates of weakly
deleterious mutations by accurately modeling levels of genic fitness variation.
We extend Santiago and Caballero’s (hereafter the SC16) model so that it can be
used to model genome-wide patterns of diversity under negative selection
processes, and develop Python software to use this theoretic approach to fit
genome-wide models of negative selection. Using forward simulations, we show
this model leads to more accurate DFE estimates under weak selection. While
this new model accommodates the possibility that a given reduction B could be
due to weaker selection, in applying our composite-likelihood method to human
population genomic data, we confirm earlier findings that strong negative
selection at a small fraction of sites fits human data across populations best.
Finally, we show that while this model solves some limitations of the classic
BGS model, simulations reveal another mysterious defect: a poor fit around
$Ns=1$. We show that this can be partially corrected by incorporating the
effects of selective interference on genic variation, but the remaining
discrepancy is due to the build up of negative linkage disequilibrium that
cannot be adequately captured by this model. We find accounting for genic
selective interference between segments fits human population genomic data
slightly better, suggesting levels of negative selection in humans are
sufficient to perturb the selection at other sites.

@@

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. We hypothesized that the higher error in this domain may be
because when our method numerically solves Equations \eqref{eq:main_eqns} for
each segment, it does not consider the reduction experienced due to selection
at other linked segments. In particular, we use a fixed drift-effective
$N=1000$ corresponding to the number of diploids in the simulations rather than
$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
for each segment at position $x$ taking into account the reduction experienced
due to selection at other segments by setting $B'(x)N$ as the population size
(see Method XXX). We find the locally-rescaled B' maps reduce the relative
error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
dashed colored lines), but does not entirely eliminate the error in the $2Ns
\approx 1$ domain.

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
when mutation rates are higher (Supplementary Figure XXX). As $Ns \to 0$, the
variance in LD inflates as expected \parencite{Ohta1969-ae,Hill1968-ue}.
Overall, the build up of negative LD is consistent our view that the
equilibrium fitness variance modeled by the SC16 theory is the additive
\emph{genic} fitness variance, which differs from the additive genetic variance
by the sum of linkage disequilibria between selected sites (i.e. $\delta_{LD} =
s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$ is the LD between sites $i$ and
$j$). However, according to theory, the reductions in diversity should be
determined by levels of additive genetic fitness variance that include the
contribution of LD (Supplementary Materials Section XXX).

While we find evidence that local-recalling reduces error in the $2Ns \approx
1$ domain, we do not employ it in our genome-wide inference method for a few
reasons. First, the reduction map

numerically solving the system of equations

it currently cannot be used during when fitting
genome-wide models.

@@

Our method takes tracks of annotated features (an ``annotation model") that are
\emph{a priori} expected to have similar fitness effects, and estimates the
distribution of fitness effects for each feature type. We consider two classes
of annotation models: (1) CADD score models, which assign a rank-order
pathogenicity score to each basepair, and (2) and more interpretable gene
feature-based models that includes protein coding regions, intron and UTRs, and
PhastCons regions. We include PhastCons regions because there is abundant
evidence of highly-conserved non-coding regions in humans under negative
selection
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
would not be included in gene-based annotation. Since our method fits a DFE for
each feature type, PhastCons regions that overlap other gene annotations (e.g.
coding sequences) must be assigned to either of the two feature types. Since
this can impact estimates, we fit both alternatives: a PhastCons Priority
model, where genic features that overlap PhastCons regions are classified as
PhastCons, and a Feature Priority model, where gene structure features take
priority and the PhastCons type catches all highly-conserved non-genic regions.
In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
Priority, and Feature Priority) to high-coverage 1000 Genome data for three
populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) reference
samples. We assess and compare our models according to how well they predict
patterns of diversity on whole chromosomes left-out during the model fitting
process. We use the metric $R_\text{LOO}^2$, which is the proportion of the
observed variance in genomic diversity at the megabase scale predicted by our
model. We experimented with a few smaller spatial scales, but our results were
consistent with previous results suggesting the human linked selection signal
fits best at the megabase scale \parencite{Murphy2022-sj}.

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
    populations (colors) fit at the megabase-scale. Round points are our B' 
    method and diamonds are the classic B. Lighter color round points are
    the out-sample $R_\text{LOO}^2$ estimates for our B' method, and arrows show the decline
    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOO}^2$ 
    were not calculated for classic B values due to computational costs). The 
    horizontal dashed lines are the $R_\text{drift}^2$ expected when the residual
    variance is given by the theoretic variance in coalescence times due to drift alone.}
  \label{fig:figure-2}
\end{figure}

Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
as well (Figure \ref{fig:figure-2}A), explaining $R^2=68.45$\% and
$R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
megabase-scale, respectively. In populations impacted by the out-of-Africa
bottleneck, the goodness-of-fit was lower across all models (e.g. 62.98\% and
59.98\% for CEU and CHB respectively in the PhastCons Priority model). The CADD
6\% model was previously found by \textcite{Murphy2022-sj} to be the
best-fitting model using classic BGS. Our PhastCons Priority model fits
marginally better than the CADD 6\% model consistently across all three
populations (Supplementary Table XXX). This is likely because the PhastCons
Priority model has more DFE parameters, and thus can fit finer-scale
differences in the DFE across its four feature types. By contrast, the CADD 6\%
track must fit patterns of diversity to conserved sites under a single DFE. The
CADD 8\% model fit slightly worse than the CADD 6\% across all populations, and
the Feature Priority model had the poorest-fit.

Since our method is built upon theory that fixes the weak selection problem of
classic BGS theory, it should in principle fit equally well when an annotation
model includes regions that are likely neutrally evolution. To test this, for
each annotation model (which are ``sparse") we fit a ``full" model that assigns
the remaining complement of the genome as a feature called ``other". This other
category is likely to predominantly neutrally evolving. Ideally, the
goodness-of-fit of model using our B' method should be invariant to whether an
annotation model is sparse or full. Indeed, we find this to be the case: both
in-sample $R^2$ and out-sample $R_\text{LOO}^2$ values are nearly identical
across full and sparse-track models (Figures \ref{fig:figure-2} A and B, round
points).

By contrast, full annotation models fit poorly when using classic BGS theory
(Figure \ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable
parameter estimates (discussed below). Additionally, when sparse annotation
models contain genomic features that are likely under weak constraint (such as
introns and UTRs), models fit worse under classic BGS theory than our B' method
(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
goodness-of-fit is nearly identical between B' and classic BGS methods. This
behavior is what we would expect given that the CADD annotation models contain
only the most pathogenic sites, which are \emph{a prior} very likely under the
strong selection domain under which B' and classic BGS theory agree.

Since our $R_\text{LOO}^2$ estimates suggest our models fit left-out
chromosomes well even though our method assumes constant demography and
homogeneous mutation rates along the genome, we wondered what $R^2$ we could
expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
and $\var_i(\varepsilon_i)$ was determined by the variance in neutral
coalescence times alone with an effective population size rescaled by $B_i'$.
We calculate this ``coalescence $R^2$" by calculating the theoretic coalescence
time variance in a megabase window with recombination and an effective
population size of $B_i' N_e$ where we use our best-fitting model's as a
plug-in estimator of $B_i'$. We note that this is a rough approximation, since
this calculation assumes constant demography and that the variance in
coalescence times under negative selection is likely lower than expected under
drift alone.

Still, we find that our out-sample $R_\text{LOO}^2$ for Yoruba individuals
($R_\text{LOO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
\approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
\approx 64$\% compared to the observed $R_\text{LOO}^2 \approx 62$\% for CEU
and $R_\text{LOO}^2 \approx 59$\% for CHB. We note that because bottlenecks
increase the variance in coalescence times beyond the level implied by the
effective population size, the disagreement between observed $R_\text{LOO}^2$
and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
that negative selection models explain the vast majority coalescence time
variation at the megabase-scale that is capable of being explained (i.e. that
is not coalescence noise).

@@

@@

@@

Our binomial likelihood models the number of different pairwise comparisons
observed per window given the total number of pairwise comparisons. The
probability is given by $\bar{\pi}(b; \Psi) = \pi_0 \bar{B}(b; \mu,
\mathbf{W})$, where bars indicate averages over some bin width, and the
reduction is

\begin{align}
    \log\left(B(x; \mu, \mathbf{W}) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{j=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g}).
\end{align}
%
Here, $\mu$ is the mutation rate and $W_{j,k(g)}$ is a DFE entry in the
$\mathbf{W}$ is the $n_s \times K$ DFE matrix over the selection grid and
features. See Supplementary Materials Sections \ref{supp:data-summary} and
\ref{supp:likelihood} for more details.

@@

\subsection*{Substitution Rate Predictions}

Our MLE fit $\hat{\Phi}$ also implies a set of corresponding substitution rate
predictions for each segment under mutation-selection-draft process. There are
two sources of variation in these predictions in this model. First, there is
the variation across the pre-specified functional categories, e.g. CDS, UTRs,
and PhastCons. Second, there is region-specific variation

>>>>

%\title{Towards a Complete Model of the Negative Selection Process in Humans}
%\title{An Evolutionary Quantitative Genetic Model of the Negative Selection Process in Humans}
\title{A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans}

@@

The continual influx of new mutations into populations is the ultimate source
of all adaptations, but the vast majority of mutations either do not affect
fitness or are deleterious. Negative selection (i.e. purifying selection) works
to eliminate such deleterious mutations from the population, which is essential
to maintaining the integrity of information in the genome needed to maintain a
lineage. Mutations under negative selection are kept at low frequencies
(Haldane?), contributing to human traits (XXX) and diseases. Moreover, the
negative selection process determines sequence conservation across phylogenetic
timescales, which is evident in genome-wide cross-species comparisons
\parencite{Siepel2005-wh}. As with most selection in the genome, the negative
selection process stochastically perturbs the allele frequencies of neighboring
linked variation. This reduces genetic variability to create a large-scale
spatial linked selection ``reduction signal" in genome-wide diversity. Mutation
and negative selection's contribution to this signal is mediated by the
deleterious mutation rate, the strength of selection in conserved segments, and
the spatial distribution of recombination rates and conserved segments. Since
genome-wide recombination maps and putatively conserved segments are available
for many species, researchers have combined these with population genetic
theory to estimate the mutation rate and selection coefficients of deleterious
mutations, as well as the overall reduction in genetic variability due to the
negative selection process.

@@

In this work, we extend another class of linked selection models that derive
from quantitative genetics to address limitations of classic background
selection theory. These models suggest that the reduction in pairwise
coalescence times at a position in the genome is determined by the
recombination map and distribution of polygenic additive fitness variance along
the chromosome \parencite{Santiago1998-bs}. This reduction is due to cumulative
effect of stochastic perturbations from selection at other sites. While these
models can theoretically accommodate polygenic fitness variation from any
source as long as its rate of change is not too rapid, we focus specifically on
a deleterious-mutations-only model from Santiago and Caballero (2016). This
model is identical to the classic BGS model when selection against deleterious
mutations is strong, but it also correctly predicts the reduction in diversity
and substitutions rates of weakly deleterious mutations by accurately modeling
levels of genic fitness variation. We extend Santiago and Caballero’s
(hereafter the SC16) model so that it can be used to model genome-wide patterns
of diversity under negative selection processes, and develop Python software to
use this theoretic approach to fit genome-wide models of negative selection.
Using forward simulations, we show this model leads to more accurate DFE
estimates under weak selection. While this new model accommodates the
possibility that a given reduction B could be due to weaker selection, in
applying our composite-likelihood method to human population genomic data, we
confirm earlier findings that strong negative selection at a small fraction of
sites fits human data across populations best. Finally, we show that while this
model solves some limitations of the classic BGS model, simulations reveal
another mysterious defect: a poor fit around $Ns=1$. We show that this can be
partially corrected by incorporating the effects of selective interference on
genic variation, but the remaining discrepancy is due to the build up of
negative linkage disequilibrium that cannot be adequately captured by this
model. We find accounting for genic selective interference between segments
fits human population genomic data slightly better, suggesting levels of
negative selection in humans are sufficient to perturb the selection at other
sites.

@@

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. We hypothesized that the higher error in this domain may be
because when our method numerically solves Equations \eqref{eq:main_eqns} for
each segment, it does not consider the reduction experienced due to selection
at other linked segments. In particular, we use a fixed drift-effective
$N=1000$ corresponding to the number of diploids in the simulations rather than
$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
for each segment at position $x$ taking into account the reduction experienced
due to selection at other segments by setting $B'(x)N$ as the population size
(see Method XXX). We find the locally-rescaled B' maps reduce the relative
error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
dashed colored lines), but does not entirely eliminate the error in the $2Ns
\approx 1$ domain. Overall, while we find evidence that local-rescaling
slightly reduces error in the $2Ns \approx 1$ domain, we do not use it during
in our genome-wide inference method. This is because to avoid circularity and
possible overfitting, the method would need to solve the equilibria equations
for all segments simultaneously, but this is currently computational
unfeasible.

Finally, we hypothesized that the remaining error is because of the build up of
negative linkage disequilibria between selected sites due to Hill--Robertson
interference \parencite{Hill1966-kd,McVean2000-bt,Comeron2007-wq}. To assess
this possibility, we calculated the sum of all linkage disequilibria in our
chromosome-wide simulations. We find negative linkage disequilibria build ups
around $2Ns \approx 1$ (Figure \ref{fig:figure-1}D, top row) and is stronger
when mutation rates are higher. As $Ns \to 0$, the variance in LD inflates as
expected \parencite{Ohta1969-ae,Hill1968-ue}. Overall, the build up of negative
LD is consistent our view that the equilibrium fitness variance modeled by the
SC16 theory is the additive \emph{genic} fitness variance, which differs from
the additive genetic variance by the sum of linkage disequilibria between
selected sites (i.e. $\delta_{LD} = s^2 \sum_{i\ne j} D_{i,j}$ where $D_{i,j}$
is the LD between sites $i$ and $j$). However, according to theory, the
reductions in diversity should be determined by levels of additive genetic
fitness variance that include the contribution of LD (Supplementary Materials
Section \ref{supp:theory}).

@@

Our composite likelihood method takes tracks of annotated features (an
``annotation model") that are \emph{a priori} expected to have similar fitness
effects, and estimates the overall mutation rate and distribution of fitness
effects for each feature type. We consider two classes of annotation models:
(1) CADD score models, which assign a rank-order pathogenicity score to each
basepair, and (2) and more interpretable gene feature-based models that
includes protein coding regions, intron and UTRs, and PhastCons regions. We
include PhastCons regions because there is abundant evidence of
highly-conserved non-coding regions in humans under negative selection
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, but
would not be included in gene-based annotation. Since our method fits a DFE for
each feature type, PhastCons regions that overlap other gene annotations (e.g.
coding sequences) must be assigned to either of the two feature types. Since
this can impact estimates, we fit both alternatives: a PhastCons Priority
model, where genic features that overlap PhastCons regions are classified as
PhastCons, and a Feature Priority model, where gene structure features take
priority and the PhastCons type catches all highly-conserved non-genic regions.
In total, we fit four annotation models (CADD 6\%, CADD 8\%, PhastCons
Priority, and Feature Priority) to high-coverage 1000 Genome data for three
populations: Yoruba (YRI), Han Chinese (CHB), and European (CEU) reference
samples. We assess and compare our models according to how well they predict
patterns of diversity on whole chromosomes left-out during the model fitting
process (e.g. leave-one-chromosome-out, LOCO). We use the metric
$R_\text{LOCO}^2$, which is the proportion of the observed variance in genomic
diversity at the megabase scale predicted by our model. We experimented with a
few smaller spatial scales, but our results were consistent with previous
results suggesting the human linked selection signal fits best at the megabase
scale \parencite{Murphy2022-sj}.

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
    populations (colors) fit at the megabase-scale. Round points are our B' 
    method and diamonds are the classic B. Lighter color round points are
    the out-sample $R_\text{LOCO}^2$ estimates for our B' method, and arrows show the decline
    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOCO}^2$ 
    were not calculated for classic B values due to computational costs). The 
    horizontal dashed lines are the $R_\text{drift}^2$ expected when the residual
    variance is given by the theoretic variance in coalescence times due to drift alone.}
  \label{fig:figure-2}
\end{figure}

Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
as well (Figure \ref{fig:figure-2}A), explaining $R^2=68.45$\% and
$R^2=68.04$\% of the out-sample variance in Yoruba pairwise diversity at the
megabase-scale, respectively. In populations impacted by the out-of-Africa
bottleneck, the goodness-of-fit was lower across all models (e.g. 62.98\% and
59.98\% for CEU and CHB respectively in the PhastCons Priority model). The CADD
6\% model was previously found by \textcite{Murphy2022-sj} to be the
best-fitting model using classic BGS. Our PhastCons Priority model fits
marginally better than the CADD 6\% model consistently across all three
populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
differences in the DFE across its four feature types. By contrast, the CADD 6\%
track must fit patterns of diversity to conserved sites under a single DFE. The
CADD 8\% model fit slightly worse than the CADD 6\% across all populations, and
the Feature Priority model had the poorest-fit.

Since our method is built upon theory that fixes the weak selection problem of
classic BGS theory, it should in principle fit equally well when an annotation
model includes regions that are under no selective constrain and neutrally
evolving. To test this, for each annotation model (which are ``sparse") we fit
a corresponding ``full" model that assigns the remaining complement of the
genome as a feature called ``other". This category contains most of the genome,
and is likely to be predominantly neutrally evolving. Ideally, the
goodness-of-fit of model using our B' method should be invariant to whether an
annotation model is sparse or full. Indeed, we find this to be the case: both
in-sample $R^2$ and out-sample $R_\text{LOCO}^2$ values are nearly identical
across full and sparse-track models (Figures \ref{fig:figure-2}A and
\ref{fig:figure-2}B, round points).

By contrast, full annotation models fit poorly when using classic BGS theory
(Figure \ref{fig:figure-2}B, diamond-shaped points) and lead to unreasonable
parameter estimates (discussed below). Additionally, when sparse annotation
models contain genomic features that are likely under weak constraint (such as
introns and UTRs), models fit worse under classic BGS theory than our B' method
(Figure \ref{fig:figure-2}A). However, among the CADD annotation models, the
goodness-of-fit is nearly identical between B' and classic BGS methods. This
behavior is what we would expect given that the CADD annotation models contain
only the most pathogenic sites, which are \emph{a priori} very likely under the
strong selection domain under which B' and classic BGS theory agree.

Since our $R_\text{LOCO}^2$ estimates suggest our models fit left-out
chromosomes well even though our method assumes constant demography and
homogeneous mutation rates along the genome, we wondered what $R^2$ we could
expect under a simple theoretic model where $\cov_i(B_i', \varepsilon_i) = 0$
and $\var_i(\varepsilon_i)$ was determined by the variance in neutral
coalescence times alone with an effective population size rescaled by $B_i'$.
We calculate this ``coalescence $R^2$" by calculating the theoretic coalescence
time variance in a megabase window with recombination and an effective
population size of $B_i' N_e$ where we use our best-fitting model's as a
plug-in estimator of $B_i'$. We note that this is a rough approximation, since
this calculation assumes constant demography and that the variance in
coalescence times under negative selection is likely lower than expected under
drift alone.

Still, we find that our out-sample $R_\text{LOCO}^2$ for the Yoruba samples
($R_\text{LOCO}^2 \approx 68$\%) is close to the theoretic $R_\text{coal}^2
\approx 67$\%. For bottlenecked out-of-Africa populations, $R_\text{coal}^2
\approx 64$\% compared to the observed $R_\text{LOCO}^2 \approx 62$\% for CEU
and $R_\text{LOCO}^2 \approx 59$\% for CHB. We note that because bottlenecks
increase the variance in coalescence times beyond the level implied by the
effective population size, the disagreement between observed $R_\text{LOCO}^2$
and the theoretic $R_\text{coal}^2$ is likely to shrink. Overall, this suggests
that negative selection models explain the vast majority coalescence time
variation at the megabase-scale that is capable of being explained (i.e. that
is not coalescence noise).

@@

Linked selection theory suggests the reduction signal is proportional to the
spatial distributions of recombination and fitness variance along the
chromosome.

@@

Each of our maximum likelihood estimates $\widehat{\Psi}$ imply other
corresponding

@@

Our binomial likelihood models the number of different pairwise comparisons
observed per window given the total number of pairwise comparisons. The
binomial probability for window $b$ is $\bar{\pi}(b; \Psi) = \pi_0 \bar{B}(b;
\mu, \mathbf{W})$, where bars indicate averages over some bin width. The free
parameters $\Psi = \{\pi_0, \mu, \mathbf{W}\}$ are the expected diversity in
the absence of selection ($\pi_0$), the mutation rate ($\mu$), and the
distribution of fitness effects for the discretized selection grid and $K$
features ($\mathbf{W}$). The reduction at position $x$ is then,

\begin{align}
    \log\left(B(x; \mu, \mathbf{W}) \right) = - \frac{1}{2} \sum_{g \in G} \sum_{j=1}^{n_s} V_g(\mu W_{j, k(g)}, s_j) Q_g^2(m_i, s_j, r_{x, g}).
\end{align}
%
See Supplementary Materials Sections \ref{supp:data-summary} and
\ref{supp:likelihood} for more details.

@@

\subsection*{Substitution Rate and Total Genic Fitness Variance Predictions}

a set of corresponding substitution rate predictions for each segment under
mutation-selection-draft process. There are two sources of variation in these
predictions in this model. First, there is the variation across the
pre-specified functional categories, e.g. CDS, UTRs, and PhastCons. Second,
there is region-specific variation

The SC16 theory models linked selection in quantitative genetics terms and
relies on simultaneously predicting the deleterious substitution rate.

### 2023-07-02 manuscript/supplementary.tex
minor ms changes

  \caption{The distribution of diversity across the genome at the megabase
  scale, with outliers trimmed. The blue line is the normal CDF, fit with MLE
parameters XXX. The inset figure is the untrimmed data, with trimmed points
shown in red.}
  \label{suppfig:trimming}
\end{figure}

The likelihood method is fit to non-overlapping binned summaries of the allele
counts matrix.

Based on exploratory data analyses, some bins were outliers and
excluded (XXX).

\begin{enumerate}
  \item \textbf{Fraction of inaccessible sites per window}.  \path{mask_inaccessible_bins_frac}
  % \item \textbf{Edge truncation}

  \item \textbf{Outliers}: Based on exploratory analysis, there were some
    regions with very high diversity. The 0.05\% right tail was excluded.

% supp data
% - npy files
% - edge trunation

@@

@@

\caption{$R^2$ and mutation rate estimates for all models. XXX note about repeats}

@@

\subsection{Derivation of Approximate Coalescence $R^2$}
\label{supp:r2-coal}

>>>>

  \caption{The distribution of diversity across the genome at the megabase
  scale, with outliers trimmed. The blue line is the best-fitting normal CDF.
  We note that this implies the distribution of megabase-scale diversity 
  is slightly fat-tailed compared to a normal distribution.
  The inset figure is the untrimmed data, with trimmed points shown in red. }
  \label{suppfig:trimming}
\end{figure}

Based on exploratory data analyses, we noticed some rare windows that were
outliers. Given that maximum likelihood estimation can be easily biased by
outliers, we took a conservative approach and removed the top 0.5\% of
high-diversity windows. The results of this can be seen in Supplementary
Figure \ref{suppfig:trimming}. We also remove 1cM from each end of the
recombination map. These window-based filters are in addition to the
accessible and neutral sequence masked described in Supplementary Sections
\ref{supp:filter} and \ref{supp:accessible}.

@@

    \label{supp:tbl-r2}

@@

\caption{$R^2$ and mutation rate estimates for all models. Note that the repeated values in the last 
rows are correct and due to rounding; see \path{notebooks/main_fits.ipnb} in the GitHub repository for 
more information.}

@@

\subsection{Derivation of Approximate Coalescence $R_\text{coal}^2$}
\label{supp:r2-coal}
